---
title: "Disruption Testing"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
---

<!-- Add the following line manually to the rendered HTML document so that IE does not block the javascript elements: -->
<!-- saved from url=(0014)about:internet --> 

```{r setup, include=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(ggplot2)
library(dplyr)
library(tidyr)
library(DT)
library(plotly)
library(knitr)
library(rgdal)
library(DBI)

base_scen = 'C:/FTOT/scenarios/quick_start/qs7_rmp_proc_dest_multi_inputs'

```

# Plotting the disruption scenarios

Total scenario cost versus disruption steps. The number of edges (roadway segments) used at each step is shown by color of the step.

```{r BC_disrupt}
disrupt_type = 'BC'

disrupt_root = paste(base_scen, disrupt_type, 'disrupt', sep = '_')

res = read.csv(file.path(disrupt_root, 'Results.csv'))

# clean up commas
res$total_cost <- as.numeric(sub('\\,', '', res$total_cost))

g1 <- ggplot(res, aes(x = disrupt_step, 
                      y = total_cost)) +
  geom_step(size = 2, color = 'grey80') +
  geom_point(size = 2, aes(color = nedge)) +
  theme_bw() +
  ggtitle('Betweeness centrality disruption of QS 7 scenario')

ggsave(plot = g1, filename = 'BC_Disruption_Cost.jpg')


ggplotly(g1)

```

## Compare evenness and network performance


```{r, read_road_data}
gdb_dir = 'C:/FTOT/scenarios/common_data/networks/Public_Intermodal_Network_2019_3.gdb'

if(!file.exists('RoadLayer.RData')){
  road = readOGR(gdb_dir, layer = 'road')
  save(list = 'road', file = 'RoadLayer.RData')
  } else {
    load('RoadLayer.RData')
  }

# First, get IDs
road_ID = unlist(lapply(road@lines, function(x) x@ID))

road_dat = data.frame(road_ID, road@data)
```

```{r, evenness_function}
evenness_func <- function(d, measure = 'capac_minus_volume_zero_floor', bin_by = NULL, n_bin = 10){
  # bin_by = 'max_edge_capacity'
  # 
  if(is.null(bin_by)){
    S = nrow(d)
    Hmax = log(S)
    
    sum_weight = sum(d[,measure], na.rm = T)
    props = d[,measure] / sum_weight 
    Hprime = -1 * sum(log(props[!is.na(props)]) * props[!is.na(props)], na.rm=T)
    Evenness = Hprime / Hmax
  }
  if(!is.null(bin_by)){
    dx = d
    dx$bins = cut(dx[,bin_by], breaks = n_bin)
    
    S = n_bin
    Hmax = log(S)
    
    props = dx %>%
      ungroup() %>%
      mutate(sum_weight = sum(get(measure), na.rm=T)) %>%
      group_by(bins) %>%
      summarize(class_sum = sum(get(measure), na.rm=T),
                props = mean(class_sum  / sum_weight)) 
    
    Hprime = -1 * sum(log(props$props[!is.na(props$props)]) * props$props[!is.na(props$props)], na.rm=T)
    Evenness = Hprime / Hmax
    
  }
  Evenness
  
}
```


```{r}
# Identify scenario name
split_path <- function(path) {
     rev(setdiff(strsplit(path,"/|\\\\")[[1]], ""))
} 

scen = split_path(disrupt_root)[1]

# read in edges from main.db
con <- dbConnect(RSQLite::SQLite(),
                 file.path(base_scen, 'main.db'))

d <- dbReadTable(con, "edges")

d = d %>%
  filter(mode == 'road')



# Get functional class

d$mode_oid = as.character(d$mode_oid)

# join road attributes to these edges
d <- d %>%
  left_join(road_dat, by = c('mode_oid' = 'road_ID')) %>%
  filter(!duplicated(mode_oid))

# 1a. Available Capacity, zero floor
e1a <-evenness_func(d, 'capac_minus_volume_zero_floor', bin_by = NULL)

# 1b. Volume-capacity ratio
e1b <- evenness_func(d, 'VCR', bin_by = NULL)

# 2. Binned by capacity

# 2a: Volume Capacity Ratio  

e2a <- evenness_func(d, 'VCR', bin_by = 'max_edge_capacity')

# 2b: Sum miles of roadway

e2b <- evenness_func(d, 'MILES', bin_by = 'max_edge_capacity')

results = data.frame(scen, e1a, e1b, e2a, e2b)

write.csv(results, file = 'Evenness_Calcs.csv', row.names = F)
```

This plots shows the same results, but now as a percent change in total scenario cost, for easier comparison with alternative scenario runs. 

```{r}
# Join with scenario cost outputs ----

res = read.csv(file.path(disrupt_root, 'Results.csv'))

# clean up commas
res$total_cost <- as.numeric(sub('\\,', '', res$total_cost))

pct_ch = res %>%
  mutate(pct_ch = 100*(total_cost - res$total_cost[1])/res$total_cost,
         scen = scen) %>%
  select(scen, disrupt_step, pct_ch)

# original version
g_i <- ggplot(res, aes(x = disrupt_step,
                      y = total_cost)) +
  geom_step(size = 2, color = 'grey80') +
  geom_point(size = 2) + #, aes(color = nedge)) +
  theme_bw() +
  ggtitle(scen) +
  xlab('Disruption Step') + ylab('Total Scenario Cost')+ 
  labs(subtitle = paste0(' 2a. Evenness based on VCR: ', round(results[results$scen == scen, 'e2a'], 3), '\n 2b. Evenness based on Miles: ', round(results[results$scen == scen, 'e2b'], 3)))

assign(paste0('gplot_', scen), g_i)

ggsave(plot = get(paste0('gplot_', scen)),
       file = "Evenness_Disruption_Fig.jpeg",
       width = 8, height = 10, dpi = 400, units = 'in')

# New plot: percent change over steps, for each 
scen_pct_ch <- left_join(pct_ch, results, by = 'scen')

scen_pct_ch2 <- scen_pct_ch %>%
  mutate(scen_lab = NA) %>%
  group_by(scen) %>%
  mutate(scen_lab = ifelse(disrupt_step == 50, scen, NA))
# 
# ggplot(scen_pct_ch,
#        aes(x = disrupt_step,
#            y = pct_ch,
#            color = e2a)) +
#   geom_line() + 
#   facet_wrap(~scen)


e2a <- ggplot(scen_pct_ch2,
              aes(x = disrupt_step,
                  y = pct_ch,
                 # color = e2a,
                  group = scen)) +
  ylab('Percent change in  total cost') +
  xlab('Disruption step') +
  geom_line(size = 1.5, alpha = 0.8, color = 'grey80') +
  geom_point(size = 2) +
  # geom_text(aes(label = scen_lab), check_overlap = F,
  #           nudge_x = 2) +
  labs(color = 'Evenness: \n Capacity-binned, VCR') +
  theme_bw() +
  ggtitle('Percent change in total scenario cost \n by capacity-binned VCR Evenness')


e2b <- ggplot(scen_pct_ch2,
              aes(x = disrupt_step,
                  y = pct_ch,
                  color = e2b,
                  group = scen)) +
  ylab('Percent change in  total cost') +
  xlab('Disruption step') +
  geom_line(size = 1.5, alpha = 0.8) +
  geom_text(aes(label = scen_lab), check_overlap = F,
            nudge_x = 2) +
  labs(color = 'Evenness: \n Capacity-binned, Miles') +
  theme_bw() +
  ggtitle('Percent change in total scenario cost \n by Evenness 2b. Capacity-binned, Miles')


ggsave(plot = e2b, file = "One_Panel_Pct_Change_Evenness_Disruption_Fig.jpeg",
       width = 7, height = 6, dpi = 400, units = 'in')

## Another one-panel fig: maximum percent change

max_ch = scen_pct_ch %>%
  group_by(scen) %>%
  summarize(max_pct_ch = max(pct_ch))

results = left_join(results, max_ch, by = 'scen')

# Wide to long
res_long = results %>%
  pivot_longer(cols = c('e1a', 'e1b', 'e2a', 'e2b'))

res_long$name = as.factor(res_long$name)
levels(res_long$name) = c('1a. Link-level, Available Capacity',
                          '1b. Link-level, VCR',
                          '2a. Capacity-binned, VCR',
                          '2b. Capacity-binned, Miles')


# ggplot(res_long, aes(x = max_pct_ch, y = value, color = scen)) +
#   geom_point(size = 2) +
#   facet_wrap(~name) +
#   xlab('Maximum Percent Change in Total Scenario Cost') +
#   ylab('Evenness Metric Value') +
#   labs(color = 'Scenario') +
#   theme_bw() +
#   ggtitle('Four evenness metrics compared to maximum percent change \n in total scenario cost')


ggsave(file = "One_Panel_Max_Pct_Change_4Evenness_Fig.jpeg",
       width = 7, height = 6, dpi = 400, units = 'in')

ggplotly(e2a)

```

## Relationship of betweeness centrality and volume

Using the edges in the optimal solution.

Both panels show the relationship between betweenness centrality and background commodity volume flow for the links used in the optimal solution for this scenario.

The left panel colors in the points by the rank in terms of volume. The top 50 links (used for the disruption test) are colored in.
The right panel colors int the points by the rank in terms of sum of betweeness centrality.


```{r bc_vol, fig.width=10}
edges <- read.csv(file.path(base_scen, 'Edges_to_Remove.csv'))

e_bc <- edges[order(edges$sum_BC, decreasing = T),] %>%
  mutate(BC_rank = order(sum_BC, decreasing = T)) %>%
  mutate(BC_rank = ifelse(BC_rank <= 50, BC_rank, NA))

e_v <- edges[order(edges$volume, decreasing = T),] %>%
  mutate(V_rank = order(volume, decreasing = T)) %>%
  mutate(V_rank = ifelse(V_rank <= 50, V_rank, NA))

e_j <- full_join(e_bc, e_v)

g2_v <- ggplot(e_j, aes(x = volume, 
                      y = sum_BC)) +
  geom_point(aes(color = V_rank), size = 3) +
  theme_bw() +
  ggtitle('Betweeness centrality sum and Volume of edges \n in optimal QS 7 scenario solution')

g2_bc <- ggplot(e_j, aes(x = volume, 
                      y = sum_BC)) +
  geom_point(aes(color = BC_rank), size = 3) +
  theme_bw() #+
  #ggtitle('Betweeness centrality sum and Volume of edges in optimal QS 7 scenario solution')

egg::ggarrange(g2_v, g2_bc, ncol = 2)

# Zoom in on just the links identified by either BC or V

top_edges = e_j %>% filter(!is.na(BC_rank) | !is.na(V_rank))

# top_edges %>% View()

```

## Table of segment links removed

Table of segment links which were removed in the disruption steps.

```{r edge_table}

datatable(e_j  %>% filter(!is.na(BC_rank) | !is.na(V_rank)) %>%
            select(BC_rank,
                   V_rank,
                   edge_id,
                   miles,
                   capacity,
                   volume,
                   sum_BC,
                   mode_oid),
          caption = "Road segements (edges) disrupted in QS 7",
          rownames = F,
          filter = 'top',
          options = list(dom = "ftp",
                         pageLength = 10)
          ) %>% formatRound(4:5, digits = 2) 

```
