---
title: "Disruption Testing"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
---


<!-- Add the following line manually to the rendered HTML document so that IE does not block the javascript elements: -->
<!-- saved from url=(0014)about:internet --> 

```{r setup, include=FALSE, warning=F, message=F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)

library(tidyverse)
library(DT)
library(plotly)
library(knitr)
library(kableExtra)
library(ggmap) # for get_stamen_map


base_scen = 'C:/FTOT/scenarios/quick_start/qs7_rmp_proc_dest_multi_inputs'

base_scen = 'C:/FTOT/scenarios/quick_start/qs7_OH'


```

# Plotting the disruption scenarios

# BC Disruption

```{r BC_disrupt}
disrupt_type = 'BC'

disrupt_root = paste(base_scen, disrupt_type, 'disrupt', sep = '_')

res = read.csv(file.path(disrupt_root, 'Results.csv'))

# clean up commas
res$total_cost <- as.numeric(sub('\\,', '', res$total_cost))

g1 <- ggplot(res, aes(x = disrupt_step, 
                      y = total_cost)) +
  geom_step(size = 2, color = 'grey80') +
  geom_point(size = 2, aes(color = nedge)) +
  theme_bw() +
  ggtitle('Betweeness centrality disruption of QS 7 scenario')

ggsave(plot = g1, filename = 'BC_Disruption_Cost.jpg')


ggplotly(g1)

```

## Volume Disruption


```{r V_disrupt}
disrupt_type = 'V'

disrupt_root = paste(base_scen, disrupt_type, 'disrupt', sep = '_')

res = read.csv(file.path(disrupt_root, 'Results.csv'))

# clean up commas
res$total_cost <- as.numeric(sub('\\,', '', res$total_cost))

g1 <- ggplot(res, aes(x = disrupt_step, 
                      y = total_cost)) +
  geom_step(size = 2, color = 'grey80') +
  geom_point(size = 2, aes(color = nedge)) +
  theme_bw() +
  ggtitle('Volume disruption of QS 7 scenario')

ggsave(plot = g1, filename = 'V_Disruption_Cost.jpg')


ggplotly(g1)

```

## Relationship of betweeness centrality and volume

Using the edges in the optimal solution

Both panels show the relationship between betweenness centrality and background commodity volume flow for the links used in the optimal solution for this scenario.

The left panel colors in the points by the rank in terms of volume. The top 50 links (used for the disruption test) are colored in.
The right panel colors int the points by the rank in terms of sum of betweeness centrality.

Of the 250 road segments used in this solution, the 


```{r bc_vol, fig.width=10}
edges <- read.csv(file.path(base_scen, 'Edges_to_Remove.csv'))

e_bc <- edges[order(edges$sum_BC, decreasing = T),] %>%
  mutate(BC_rank = order(sum_BC, decreasing = T)) %>%
  mutate(BC_rank = ifelse(BC_rank <= 50, BC_rank, NA))

e_v <- edges[order(edges$volume, decreasing = T),] %>%
  mutate(V_rank = order(volume, decreasing = T)) %>%
  mutate(V_rank = ifelse(V_rank <= 50, V_rank, NA))

e_j <- full_join(e_bc, e_v)

g2_v <- ggplot(e_j, aes(x = volume, 
                      y = sum_BC)) +
  geom_point(aes(color = V_rank), size = 3) +
  theme_bw() +
  ggtitle('Betweeness centrality sum and Volume of edges \n in optimal QS 7 scenario solution')

g2_bc <- ggplot(e_j, aes(x = volume, 
                      y = sum_BC)) +
  geom_point(aes(color = BC_rank), size = 3) +
  theme_bw() #+
  #ggtitle('Betweeness centrality sum and Volume of edges in optimal QS 7 scenario solution')

egg::ggarrange(g2_v, g2_bc, ncol = 2)

# Zoom in on just the links identified by either BC or V

top_edges = e_j %>% filter(!is.na(BC_rank) | !is.na(V_rank))

# top_edges %>% View()

```

Table of segment links which were removed in the disruption steps.


```{r edge_table}

datatable(e_j  %>% filter(!is.na(BC_rank) | !is.na(V_rank)) %>%
            select(BC_rank,
                   V_rank,
                   edge_id,
                   miles,
                   capacity,
                   volume,
                   sum_BC,
                   mode_oid),
          caption = "Road segements (edges) disrupted in QS 7",
          rownames = F,
          filter = 'top',
          options = list(dom = "ftp",
                         pageLength = 10)
          ) %>% formatRound(4:5, digits = 2) # %>%
#          formatStyle(1, background = styleEqual(max(tabl[,1]), 'lightgreen'))%>%
          

```
